### 2017-03-12 doc/Submission rev. 2/response.tex
Autocommit

\documentclass[twocolumn,5p]{elsarticle}
\pdfoutput=1
\usepackage{hyperref}
\usepackage{tikz}
\def\firstcircle{(90:1.75cm) circle (2.5cm)}
\def\secondcircle{(210:1.75cm) circle (2.5cm)}
\def\thirdcircle{(330:1.75cm) circle (2.5cm)}
\usepackage{comment}
\usepackage{framed,graphicx}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{amsmath}
\usepackage{bigstrut}
\usepackage{color}
\usepackage{graphics}
\usepackage{eqparbox}
\usepackage{graphics}
\usepackage{colortbl}
%\usepackage{times}
\usepackage{mathptmx} \usepackage[scaled=.90]{helvet} \usepackage{courier}
\usepackage{balance}
\usepackage{picture}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage[export]{adjustbox}
\renewcommand{\footnotesize}{\scriptsize}
\definecolor{lightgray}{gray}{0.8}
\definecolor{darkgray}{gray}{0.6}
\definecolor{lavenderpink}{rgb}{0.98, 0.68, 0.82}
\definecolor{celadon}{rgb}{0.67, 0.88, 0.69}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
%%% graph
\newcommand{\crule}[3][darkgray]{\textcolor{#1}{\rule{#2}{#3}}}

\newcommand{\quart}[4]{\begin{picture}(80,4)%1
	{\color{black}\put(#3,2){\circle*{4}}\put(#1,2){\line(1,0){#2}}}\end{picture}}

\definecolor{Gray}{gray}{0.95}
\definecolor{LightGray}{gray}{0.975}

\newcommand{\rahul}[1]{{\color{steel}{#1}}}
\newcommand{\menzies}[1]{{\color{magenta}{#1}}}
\newcommand{\lucas}[1]{{\color{red}{#1}}}

%% timm tricks
\newcommand{\bi}{\begin{itemize}[leftmargin=0.4cm]}
	\newcommand{\ei}{\end{itemize}}
\newcommand{\be}{\begin{enumerate}}
	\newcommand{\ee}{\end{enumerate}}
\newcommand{\tion}[1]{\S#1}
\newcommand{\fig}[1]{Figure~#1}
\newcommand{\tab}[1]{Table~\ref{tab:#1}}
\newcommand{\eq}[1]{Equation~\ref{eq:#1}}
\newcommand{\review}[1]{\noindent\textit{#1\\}}

\usepackage[shortlabels]{enumitem}
\usepackage{url}

\definecolor{steel}{rgb}{.11, .11, .7}
\definecolor{Gray}{rgb}{0.88,1,1}
\definecolor{Gray}{gray}{0.85}
% \usepackage[svgnames]{xcolor}
\usepackage[framed]{ntheorem}
\usepackage{framed}
\usepackage{tikz}
\usetikzlibrary{shadows}
%\newtheorem{Lesson}{Lesson}
\theoremclass{Lesson}
\theoremstyle{break}

% inner sep=10pt,
\tikzstyle{thmbox} = [rectangle, rounded corners, draw=black,
fill=Gray!40,  drop shadow={fill=black, opacity=1}]
\newcommand\thmbox[1]{%
	\noindent\begin{tikzpicture}%
	\node [thmbox] (box){%
		\begin{minipage}{.94\textwidth}%
		\vspace{-2mm}#1\vspace{-3mm}%
		\end{minipage}%
	};%
	\end{tikzpicture}}

\let\theoremframecommand\thmbox
\newshadedtheorem{lesson}{Result}

\begin{document}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% RESPONSE TO REVIEWERS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\pagebreak
\newpage
\appendix
\renewcommand*{\thesection}{\Alph{section}}
\nobalance
\section*{RESPONSE TO REVIEWERS}
\section*{Comments from the editor}
\review{The reviewers have finished to  evaluate your work. 
	As you 
	can observe, they raised some concerns that deserve your attention~$\ldots$}

\noindent Thank you very much indeed for your comments.\\

\review{ $\ldots$  It is important that you take into account their 
	comments and 
	evolve the paper accordingly. Please,  evolve the paper to make clear the 
	defect model, hypothesis, limitations and readability}

We offer clarifications to all the issues highlighted by the reviewers. In 
addition to these, we have made several key modifications to our paper. We 
have changed the title of our paper to reflect the changes we have made. In 
order to assist reviewers in tracking our changes, we have ensured all 
new additions are {\color{steel} highlighted in blue}. 
A brief summary of our changes are listed below:
\be
\item At the suggestion of Reviewers 1 and 2, we have included a section on 
the choice of datasets. Additionally, we have greatly expanded the section 
on defect prediction to highlight its importance and its relevance to our 
work.
\item To address comments by Reviewer~\#1 on how this work fits in with the 
body of knowledge, we have compiled a new section (see~\tion{2.1}) on the 
previous attempts to address similar issues. 
\item Further, at the suggestion of Reviewer~\#1, we have included a 
section remarking on the key limitations of our work. It now appears as a 
new section (see \tion{7}).
\item Also, as a response to Reviewer~\#1's comments, we have modified the 
future work section to reflect the implications of our findings on 
\textit{technical debt}. 
\item We acknowledge Reviewer~\#2's comment that our initial submission 
lacks a formal introduction of software ``defects" and its relevance to 
code reorganization. In order to address this, we have attempted to 
clarify 
the interpretation of defects in our work. As for other issues related to 
defects, we have added a new section on the choice of datasets 
(see~\tion{3}).
\item Also, keeping in mind the suggestions of Reviewer~\#2, we have 
rewritten the research questions (especially, RQ1 and RQ5).
\item Additionally, at the suggestion of Reviewer~\#2, we clarify our 
descriptions of several figures included in this paper. Specifically, we 
expand on the importance of tuning which relates to \fig{5} (previously 
Figure 4). We also expand on the significance of \fig{9} 
(previously 
Figure 8).
\item Finally, we have evolved our paper in the most appropriate areas in 
response to all the major and minor issues raised by both Reviewer~\#1 and 
Reviewer~\#2.
\ee

\section*{Response to Reviewer~\#1}

Thank you for your detailed comments. Our new draft contains many 
corrections, clarifications, and extensions as directed by your feedback.
All our changes are marked in the body and appears in {\color{steel} 
	blue}.\\

\review{I missed a section discussing possible implications for 
	researchers 
	and practitioners. How the results of this work contribute to future 
	research activities in the area? How practitioners can benefit from the 
	results achieved in the paper?}

You're quite correct, there was no future work section in the paper. The
following text
has been included in the revised submission and appears in~\tion{7} 
(see {\color{steel} Reviewer 1a}).

\begin{quote}``Our research shows that it is potentially na\"{\i}ve to explore 
	thresholds in static code attributes in isolation to each other. 
	This work clearly demonstrates how changing one necessitates changing 
	other associated metrics. So, for future work, we recommend that 
	researchers and practitioners look for tools that recommend changes to sets 
	of code changes. When exploring candidate technologies, apart from XTREE, 
	researchers may potentially use: (1) Association rule learning; (2) 
	Thresholds generated across synthetic dimensions (eg. PCA); and (3) 
	Techniques that cluster data and look for deltas between them. (Note: we 
	offer XTREE as a possible example of this point).''

	``Additionally, we plan on extending our work by exploring scalable 
	solutions 
	to achieve similar results in much larger datasets. After this, we shall 
	look at applications beyond that of measuring static 
	code attributes. For example, as an initial attempt, we have been looking 
	at sentiment analysis in stack overflow exchanges to learn dialog pattern 
	that most select for relevant entries.''
\end{quote}

\review{I also missed a section presenting the main limitations of 
	the 
	work. Although some words have been said about it, I believe it is 
	important to have a more structured discussion on a specific section.}

You are quite correct that we have missed a section highlighting the key 
limitations of our work. In the revised paper, we have dedicated a section 
to discussing the 
limitations and future work (see {\color{steel} Reviewer1b} in 
~\tion{7}).\\

\review{It would be interesting to have further discussion at the 
	end 
	indicating how this work complements the body of knowledge about code 
	smells. Do 
	reached results indicate a similar direction? contradictory? Does This work 
	somehow confirms the results of other studies or points to new 
	directions?}

Thank you for your comment, we admit the last version of the paper has 
an inadequate discussion on this matter. This revised version now includes an 
expanded prior work section which places greater emphasis on these issues 
(see {\color{steel} Reviewer1c} in 
\tion{2.1}).

As to your specific comment that this be included in the 
end, we feel our response to this recommendation is an important note on 
the state-of-the-art in this area. We have therefore added it towards the 
beginning of the paper instead, so it may premise the rest of our work.\\

\review{$\ldots$ this work evaluated XTREE, a framework to 
	evaluate whether a code reorganization is likely to have a perceivable 
	benefit in terms of defect-proneness. It is fine to have this scope 
	(defect-proneness) in terms of evaluation. However, it would be nice to 
	have a discussion on why this characteristic (defect-proneness) is good to 
	support the decision on code reorganization and/or how it could complement 
	other characteristics.}

You are quite right in pointing out that we need a better description of 
defect proneness. We have now added a section on defect-proneness and its 
relevance to our paper (please see content marked with {\color{steel} 
	Reviewer1d} in~\tion{3}).\\

\review{$\ldots$ nowadays we are seeing a growing discussion on 
	technical debt, 
	that brings financial aspects (interest, principal, debt…) to the 
	decision-making process on development activities. Could we use XTREE as a 
	strategy to prioritize the payment of technical debt items?}

Thank you for that suggestion. We do think XTREE can be used as a 
recommendation system to prioritize technical debt. In our revised paper, 
we expanded the introduction to highlight this possibility (see 
content prefixed with {\color{steel}Reviewer1e} in \tion{7}).\\

\review{Finally, I have one last question: is there any reason for 
	do not 
	use the GQM template to state the goal of the study and also to define null 
	and alternative hypotheses for the research questions?}

We can certainly do that if you wish. As to why we {\em didn't} do it,
this paper is clearly MSR-conference-style and not
ESEM-conference-style (GQM is less uses at MSR than ESEM).

To whether or not that is a ``good thing'' or not, we cannot say. What
we would say is that we think our use of stand-out boxes for research
findings makes our results easy to access within this paper.

That said, if the reviewer thinks this paper needs further structuring
to clarify our message then, of course, we would follow that direction.

%	\texttt{XXX}

\section*{Response to Reviewer~\#2}

Thank you for your detailed review. We have made several 
corrections, clarifications, and revisions as directed by your comments, 
this new draft is significantly improved thanks to your feedback.

\noindent All our 
changes are marked in the body and appears in 
{\color{steel} blue}.\\

\review{The introduction talks about Fowler’s bad smells and 
	refers to 
	literature on the (often contradictory) effects of such smells. And Figure 
	1 shows a list of bad smells. However, without making it clear, the paper 
	continues to talk about smells but the real topic is not the smells but the 
	metrics that are used to identify various smells. What’s the connection 
	between whether a smell is bad or not and the threshold values of the 
	various code metrics?}

Thank you for this comment. At your suggestion, we have significantly 
expanded \tion{3} in order to discuss the use of static code metrics 
with regards to code smells.\\

\review{I also find it very difficult to understand the conceptual model of 
	defects in this paper~\ldots}

We do apologize for the inconvenience.
Your remark clearly indicates that we need to further clarify our use
of the term.
So at your suggestion, we have 
dedicated a section to discussing the choice of dataset in \tion{3} 
(please see content marked with 
{\color{steel} Reviewer2a}).\\

\review{\ldots~The concept of “number of defects (defect counts)” 
	is fully mixed with the concept of “percent of classes with 
	defects/defective modules/ presence or absence of defects”. What does it 
	then mean to “reduce defects in our data sets”? Is it the total number of 
	defects in the data set (system?) or the number of “infected modules” in 
	the data set?}

We have clarified this in \tion{3}. Briefly, with respect to the 
datasets used in this study, when we say, ``reduce defects in our data 
sets,'' we refer to 
reorganizing code such that our secondary oracle reports
that the number of infected modules (i.e., classes with 
$\texttt{nDefects}>0$) is reduced to having \textit{no defects}.
For notes on our {\em primary} and {\em secondary} oracle, see page2, 
column1, of this paper.

In response to your  comment, we have expanded \tion{3}, to 
better explain this in much greater detail (please see content marked with 
{\color{steel}Reviewer2b} in \tion{3}). \\

\review{ Section 2 discusses the idea of why not just ask developers
	about the effect of code smells when the research literature is
	contradictory. The authors argue that practitioners
	should not be expected resolve contradictions. My point is that it is
	so obvious that most of Section 2 should be deleted. Figure 1 as part
	of a related work section may remain.}

This is an interesting comment since it is the {\em opposite} of the feedback 
we got on a earlier draft of this paper. In fact, we had
a certain senior SE person (who the reviewer might actually know....) demanding 
``why not just use best practices from the literature?''.
That comment lead to the Fig1 survey.

That said, we quite take your point that our current 
section~\tion{2.3} is far 
too long. For this draft, prompted by
your comment, we have cut its length in half.\\

\review{\ldots~How does a log of defects look 
	like? What kind of information is recorded? ...  In case 
	four new modules are developed, is there then also a log 
	history that shows how the total number of modules in the system correlates 
	with the total number of defects? Are we talking about total number of 
	defects or only the number of “infected modules”?}

To answer these issues, we have provided a sample defect data set in 
\fig{3}. The example shows a few rows of data from Ant 1.3 
containing: (1) Module 
name; (2) 20 CK-Metrics; and (3) Defects (as a count of raw defects and a 
boolean variable). Note that \fig{5} lists all the remaining datasets and 
they all have been formated in the same way. \\

\review{From RQ1: ``This code reorganization will start with some initial 
	code base that is changed to a new code base. For example, if the bad smell 
	is loc $>$ 100 and a code module has 500 lines of code, we reason 
	optimistically that we can change that code metric to 100. Using the 
	secondary verification oracle, we then predict the number of defects in 
	new.'' What is the idea here?}

Thank you for the question. Your observation regarding this --- \emph{Is 
	the point that a log 
	history of defects has shown that modules 
	with more than 100 loc have more defects (per lines of code?) than smaller 
	modules, and then the action is to reduce the size of that module?} --- is 
accurate.  We have now modified the content in corresponding section to 
further explain this in \textbf{RQ1} (please see content marked with 
{\color{steel} Reviewer2c} in \tion{5}).\\

\review{RQ1 is  ``which of the methods is most accurate?'' without 
	presenting 
	the methods. They are just referred to in three other papers that probably 
	describe them. What are your selection criteria for comparing the tool (or 
	framework or system as you also call it) XTREE with exactly these methods?}

Thank you for your comment. We have fixed the description of \textbf{RQ1} 
to be more specific (please see content marked with 
{\color{steel} Reviewer2d} in \tion{1.1}). As to selecting primary change 
oracles, we undertook a survey of literature to identify commonly used 
methods to do this. We found three other techniques: CD from a previous 
work, VARL based thresholds~\cite{Shatnawi10}, and Statistical 
thresholding~\cite{Alves2010}.
\\

\review{It is stated that this paper reports a case study. I know that within 
the SW community is its common to use the term case study to denote just a 
demonstration of an example performed by the researchers. But in more mature 
disciplines, which we should aim to become a member of, “case study” has a 
certain meaning (e.g. Yin 2003). It would mean an evaluation in a real software 
development context. This is note the case here. See also the work by Runeson 
and Host 2009 on case studies in software engineering.}

Thank you for that comment. We can show that our definition of
``case study'' is aligned with how Per Runeson uses the term.
In 2013, we were  working on revising a data mining paper for TSE paper
where Runeson was the associate editor. He pushed us hard for better 
terminilogy and definitions, including our term ``case study''. To address his 
concerns, we wrote the following
notes about ``case studies'' and data mining papers. Note that
he accepted the point offered below and accepted the paper.

\begin{quote}

	As we read standard texts on empirical software engineering
	(e.g. Runeson \& Host\footnote{P. Runeson and M. Host, ``Guidelines for 
	conducting and reporting
		case study research in software engineering,'' Empirical Software
		Engineering, vol. 14, pp. 131–164, 2009} ) the term ``case study'' 
		often 
	refers to some act which includes observing or changing
	the conditions under which humans perform some software
	engineering action.

	In the age of data mining and modelbased
	optimization this definition should be further extended
	to include case studies exploring how data collected from
	different projects can be generalized across multiple projects.
	As Cruzes et al. comment:
	\begin{quote}
		``Choosing a suitable method for synthesizing evidence
		across a set of case studies is not straightforward...
		limited access to raw data may limit the ability
		to fully understand and synthesize studies.\footnote{P. R. M. H. 
		Daniela Cruzes, Tore Dyba, ``Case studies synthesis:
			Brief experience and challenges for the future,'' in ESEM’11: the
			International Symposium on Empirical Software Engineering and 
			Measurement,
			2011.}''
	\end{quote}
\end{quote}
Note that the above quote from Cruzes et al. is almost a summary of the
specific goals of this paper: given limited amounts of software
project data, how can we rank bad smells such as to avoid making unecessary, or 
bad, changes.
\\
%\clearpage

\review{Regarding case study, what kind of work remains before the proposed 
tool could be used by practitioners?}

We need to address a scale-up problem (which is the current focus of Mr. 
Krishna's Ph.D. work) and once that is done,
we will make this a Python package freely available to the world on Windows,
Linux, and Mac machines via the simple command `\texttt{pip install xtree}'.
After that is installed, the code would search a directory of project data to 
automatically generate the figures of this paper.\\

\review{What’s the difference between tool, framework, and method in this 
	paper?}

Thank you for pointing that out. We offer the following definitions: (1) 
Framework: A basic structure underlying recommending reorganization. It 
comprises of a primary change oracle and a secondary verification oracle; 
(2) Method: Refers to how the primary change oracle offers these 
recommendation. 
In this work it may be based on Cluster Deltas (XTREE) or threshold based 
(Shatnawi and Alves); and (3) Tool: Comprises of any one primary change 
oracle (XTREE) and a secondary verification oracle (Tuned Random Forest).

We may have used these interchangeably in the first draft of our paper. 
This has been fixed in this revision.\\

\review{How does this work related to the work by Arcelli Fontana, in 
	particular: Arcelli Fontana, F., Mantyla, M.V., Zanoni, M. et al. Comparing 
	and experimenting machine learning techniques for code smell detection, 
	Empir Software Eng (2016) 21: 1143. doi:10.1007/s10664-015-9378-4?}

Thank you for the reference. Our work differs from this specific paper in 
one 
key aspect --- we attempt to assist reorganization of code and prevent 
needless reorganization while they focus on 
detecting code smells using machine learning techniques. Our work is much 
closer to other work by Arcelli Fontana~\cite{font1}. We discuss this 
in~\tion{2.1}. \\

\review{Improving on one metric may cause degradation 
	on another one. This issue could have been much more spelt out with good 
	examples in the paper. It's stated that XTREE “recommends” 
	(or “suggests”) increasing coupling while reducing LOC of a module. Sure, 
	if you split a module into several smaller modules, or move code from one 
	large module to other, smaller modules, usually the overall coupling will 
	increase.  But isn't that an avoidable consequence that will occur 
	implicitly in the process of module splitting or moving code between 
	modules? The way it’s formulated now gives the impression that the 
	programmers should follow the recommendation of increasing the coupling, 
	that is, as if it’s a conscious action. Or am I missing the point?}

You are quite right, we have now rewritten associated text with 
\textbf{RQ5} in \tion{6.5} to better explain 
the underlying connectedness of the metrics ({\color{steel} Reviewer2e} in 
\tion{6.5}).

To summarize, we were trying to convey the following information with this 
specific example:

\be

\item
Attempting to reduce certain code metrics (like LOC) necessitates 
increasing other metrics. Any recommendation that fails to respect these 
``co-changes'' may not the practical to follow.
\item  XTREE is aware of 
the relationship between metrics and always recommends changes to 
conjunctions of metrics. These changes are pragmatic in that XTREE 
sometimes suggests the same metric be either increased or decreased based 
on the dataset. As an example, consider \fig{8} in Ant coupling 
between objects (CBO) has to be reduced while in Lucene CBO has to be 
increased.
\ee

Developers are now aware of these interconnected changes. They can proceed 
to make these changes because of the historical evidence of its 
effectiveness.\\

\review{In Section 4: ``It can be difficult to judge the effects of 
	removing 
	bad smells. Code that is reorganized cannot be assessed just by a rerun of 
	the test suite since such reorganizations may not change the system 
	behavior (e.g., refactorings).'' Isn't the point about removing bad smells 
	to improve the code without changing the behavior (refactorings)? Why 
	cannot a test suite be run before and after if the behavior is not changed?}

It is true that refactoring improves the code by performing behavior 
preserving modifications. Evaluating this code with test suites that are 
designed to 
identify faults would not work for three reasons: (1) Test suites may not 
be 
designed to identify code smells, they only report pass/fail based on 
whether or not a specific module runs. It is entirely possible that a test 
case may pass for code that contains one or more code smells; (2) While 
smells are certainly symptomatic of design flaws, not all smells cause the 
system to fail in manner suitable for identification by test cases; and (3) 
It make take a significant amount of effort to write new test cases that 
identify bad smells, especially for stable software.

Finally, in our work we tackle reorganization which subsumes refactoring. 
Since not all reorganization efforts comprise of refactoring, that makes it 
difficult to assert when test case will pass or fail. Thus, using test 
suites is not a reliable option.

At your suggestion, we have now added this in \tion{5.2} (please see text 
marked {\color{steel} Reviewer2f}).\\

\balance

\review{Figure 4 shows the effect of tuning, and the authors write: ``The 
	rows marked with a * in Figure 4 show data sets whose performance was 
	improved remarkably  by  these  techniques. For example, in poi, the recall 
	increased by 4\% while the false alarm rate dropped by 21\%.'' This is a 
	good example of researcher bias. Your example is the most favorable example 
	from your point of view. It is one of only two of the eight data sets that 
	improved on both pd and pf.}

You are right \ldots ~there are only a few cases where we see an 
improvement 
in both recall and false alarm at the same time (poi, synapse and log4j). 
But, we don't think is due to a researcher bias. Instead, this due to the 
specifics of tuning and the goals used for this. We would like to draw your 
attention to two other kinds of improvements as 
a result of tuning:

\be

\item Cases where there is a significant 
improvement in one measure (recall or false alarm) with no deterioration in 
the other measure (Ivy, Jedit, Lucene, and Xalan); and

\item Cases where there is a significantly large improvement in one measure 
with acceptable deterioration in the other metric (Poi and Ant).

\ee

\noindent We attribute these to the nature of multiobjective search 
performed by 
differential evolution. As highlighted by Fu et 
al.~\cite{fu:ase15} (page 2, \S2.2), defining objectives for tuning is a 
notoriously 
difficult task. They find that tuning for recall or false alarms 
independently can lead to some undesirable results. Tuning on recall yields 
oracles that have near 100\% recall but with almost 100\% false alarm. 
Similarly, tuning for only false alarm results in oracles that have 0\% 
false alarm with 0\% recall. We take their advice and tune on metrics that 
combine both precision and recall, thereby turning this into a 
multiobjective search. Results in \fig{5} (previously Figure 4) show recall 
and false alarms have reached a ``trade-off'' with maximum improvement, 
such 
is the nature of multiobjective differential evolution.

Thank you for your comments, we ought to have had a formal discussion of 
these in the paper. At your suggestion, these observations and reasons for 
the same have now been dealt with in much greater detail in the current 
version on the paper (see text marked with {\color{steel} Reviewer2g} in 
\tion{5.2.1})
\\

\review{The authors state that if there are no historical records of 
	defects, the results of this paper can be used as a guide (which results?). 
	It is referred to Table 8 in the abstract but there is no Table 8. Is it 
	meant to be Figure 8? In case, it’s very hard to understand how that figure 
	could be used.}

A good catch ... there seems to be a typo, Table 8 has now 
been replaced by \fig{9} in the abstract. Regarding the 
understandability 
of~\fig{9}, you are quite right in that we haven't provided 
sufficient information explaining this. As per your suggestion, we have now 
expanded the section that refers to this 
figure (see text prefixed with {\color{steel} Reviewer2h} in 
\tion{6.5}) to better explain how the 
conclusions 
from this figure may be used when historical records are absent.
	\balance
\bibliographystyle{unsrt}
\bibliography{References}

\end{document}